New seismological constraints on the available parameter space of macroscopic dark matter

### 2016-09-07 LaTeX/Article_Draft.tex
Added Heat equation material
There is a key consistency check that appears to have an issue, that
may alter some of the results -- David and I plan to work on this
together later today.

The choice of $p_0 = 10^8\,\te{Pa}$ does significantly impact our result; ultimately the measured displacements we calculate will be inversely proportional to $p_0$. Our choice of $p_0$, however, is well below the overpressure corresponding to the boundary between linear and non-linear elasticity in the Moon. As the ambient pressure on a sample of rock increases, so too does $p_0$ \cite{mair2002influence}\cite{shimada1993lithosphere}\cite{ord1991deformation}. We have taken $p_0$ to be a factor of $3$ below that of granite with a modest $50\te{ MPa}$ overpressure \cite{lockner200232}. In reality, $p_0$ is probably, on average, several orders of magnitude larger than we claim it to be because the ambient pressure in the Moon is on the order of $\te{GPa}$ rather than $\te{MPa}$, but no measurements, to our knowledge, probe the elastic behavior of rock at these high pressures. It is also important to notice that this choice of $p_0$ corresponds to $r_0$ on the order of $1\te{ km}$, which well exceeds the regime of non-linearity we would have expected based on the solutions to the heat equation. 
\section{Seismic Wave Propagation}
The velocity of $p$-waves as a function of distance from the center $r$ of the Earth and Moon (and presumably other spherical rocky celestial bodies) is of the form $v = a^2 - b^2 r^2$ \cite{garcia2011very}\cite{dziewonski1981preliminary}. Using Snell's law, we obtain the differential equations for the trajectory of $p$-wave rays within such bodies
\begin{align}
v^2 \notag&= \dot r^2 + \f{p_\te{ray}^2 v^4}{r^2}\,,\\
0 &=\dot\theta\pm\f{p_\te{ray} v^2}{r^2}\,.
\end{align}
where $r$ is the radial coordinate of the ray, $\theta$ is the polar angle of the ray measured from the center of the body, and $p_\te{ray}$ is the ray parameter, which is fixed along a ray trajectory. These equations can be integrated. In the case that $v = a^2 - b^2 r^2$, we obtain for some constants $q$ and $\theta_0$.
\begin{align}
\notag&r(t) = \f{a\sqrt{(q^2e^{2abt}-b)^2 + 4a^2b^4 p_\te{ray}}}{b\sqrt{(qe^{2abt}+b)^2 + 4a^2b^4 p_\te{ray}}}\,,\\
&\tan(\theta(t) - \theta_0) = \f{qe^{4abt}-b^2}{4ab^3p_\te{ray}}+abp_\te{ray}\,.
\end{align}
\textit{A priori} these are geodesics on the Poincar\'e disc of radius $\f ab$, which is easily seen from the form of $v$.

>>>>

The choice of $p_0 = 10^8\,\te{Pa}$ does significantly impact our result; ultimately the measured displacements we calculate will be inversely proportional to $p_0$. Our choice of $p_0$, however, is well below the overpressure corresponding to the boundary between linear and non-linear elasticity in the Moon. As the ambient pressure on a sample of rock increases, so too does $p_0$ \cite{mair2002influence}\cite{shimada1993lithosphere}\cite{ord1991deformation}. We have taken $p_0$ to be a factor of $3$ below that of granite with a modest $50\te{ MPa}$ overpressure \cite{lockner200232}. In reality, $p_0$ is probably, on average, several orders of magnitude larger than we claim it to be because the ambient pressure in the Moon is on the order of $\te{GPa}$ rather than $\te{MPa}$, but no measurements, to our knowledge, probe the elastic behavior of rock at these high pressures. It is also important to notice that this choice of $p_0$ corresponds to $r_0$ on the order of $1\te{ km}$, which well exceeds the regime of non-linearity we would have expected based on the solutions to the heat equation.

\subsection{Heat Equation solution analysis}
We consider only the melting or vaporization that occurs prior to the onset of the deposited energy serving as a source for seismic waves, since the initiation of seismic wave propagation complicates assessing the amount of energy dissipated in melting (and vaporizing) the impacted material and its surroundings.  To model an instantaneous energy deposition along the z-axis, we use a delta function, and set the solution of the heat equation equal to a fixed temperature (of melting).  This yields an expression for the radius, as a function of time, of the impacted material and its surroundings that has reached its melting point temperature, $r_\te{melt}(t)$.  Differentiating $r_\te{melt}(t)$ with respect to time, setting the result equal to the bulk speed of sound and solving for t yields the time at which the propagation of the "melt point temperature zone" falls to the speed of sound.  Plugging this time into $r_\te{melt}(t)$ yields $r_\te{fast melt}$: the extent of the impacted material and its surroundings that is heated to its melting point more rapidly than bulk sound speed. Replacing the melting point temperature with the vaporization point temperature in the above demonstrates that $r_\te{fast melt}$ is essentially equal to $r_\te{fast vaporize}$ for all Macro radii.  This result indicates that for macros with radii greater than ***$10^{-9} m$***(pending resolution of consistency check issue, and minor additional refinements), $r_\te{fast vaporize}$ is of the same order as $r_\te{Macro}$.
The extreme temperatures generated by the impact (approximately $10^{15} K$, again pending resolution of consistency check issue that arose in computing this value) will convert the vaporized material to plasma; however, this conversion is a second order phase transition, and so does not require a latent heat input to occur.  Thus, the energy to transition from vapor to plasma could conceivably transfer back into seismic waves (as the particles slow down, recombine as vapor, condense to liquid, and then solidify). Similarly, the energy required to raise the temperature of the impacted material and its surroundings to their melting and vaporization points could subsequently source seismic waves: temperature changes in the column of heated material result in expansions  and contractions that lead to a column of ovepressure/underpressure. Hence, to model the energy dissipated both in raising the temperature of the impacted material and its surroundings, and in effecting a transition to plasma are complicated by the concurrent energy transfer to seismic waves.

\section{Seismic Wave Propagation}
The velocity of $p$-waves as a function of distance from the center $r$ of the Earth and Moon (and presumably other spherical rocky celestial bodies) is of the form $v = a^2 - b^2 r^2$ \cite{garcia2011very}\cite{dziewonski1981preliminary}. Using Snell's law, we obtain the differential equations for the trajectory of $p$-wave rays within such bodies
\begin{align}
v^2 \notag&= \dot r^2 + \f{p_\te{ray}^2 v^4}{r^2}\,,\\
0 &=\dot\theta\pm\f{p_\te{ray} v^2}{r^2}\,.
\end{align}
where $r$ is the radial coordinate of the ray, $\theta$ is the polar angle of the ray measured from the center of the body, and $p_\te{ray}$ is the ray parameter, which is fixed along a ray trajectory. These equations can be integrated. In the case that $v = a^2 - b^2 r^2$, we obtain for some constants $q$ and $\theta_0$.
\begin{align}
\notag&r(t) = \f{a\sqrt{(q^2e^{2abt}-b)^2 + 4a^2b^4 p_\te{ray}}}{b\sqrt{(qe^{2abt}+b)^2 + 4a^2b^4 p_\te{ray}}}\,,\\
&\tan(\theta(t) - \theta_0) = \f{qe^{4abt}-b^2}{4ab^3p_\te{ray}}+abp_\te{ray}\,.
\end{align}
\textit{A priori} these are geodesics on the Poincar\'e disc of radius $\f ab$, which is easily seen from the form of $v$.
